---
title: "ACDIP Population Project"
author: "Kenji Tomari"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(RANN)
library(sf)
library(tidyverse)
library(stringr)
library(readxl)
library(VIM)
library(osmdata)
library(readxl)
library(foreign)  # open dbf
library(stringr)
```

# ACDIP Data

## Data Cleaning Part I

```{r eval=F}
aldeas <- read_csv("data_acdip/Comunidades de ACDIP (version2) - Datos Aldeas.csv",
                   col_types=cols(.default = "c"),
                   na=c("na", "NA", "")
                   )

# remove first two rows with dummy data
aldeas <- aldeas %>%
  filter(!str_detect(`Nombre de Aldea`, "ejemplo"))  %>%
  filter(!is.na(`Nombre de Aldea`))

aldeas <- aldeas %>%
  separate(col=Coordenadas,
           into=c("lat", "long"),
           sep=";",
           extra="drop")  # to ignore the last semi-colon

# which rows seem to have insufficient digits in coords?
aldeas %>%
  filter(str_detect(lat, "\\.\\d{1,4}$") | str_detect(long, "\\.\\d{1,4}$") )
```

## Data Cleaning Part II

Spatialize Data.

```{r eval=F}
# spatialize (original data is from android app)
aldeas <- st_as_sf(aldeas,
         agr="identity",
         coords=c("long", "lat"),
         crs=4326)

# get crs for guatemala
gt_crs <- st_crs(st_read(
  "data_official/lugares_poblados_gtm/lugares_poblados_gtm.shp"))

aldeas <- st_transform(aldeas, gt_crs)
```

## Save/Read in GPKG

```{r eval=F}
st_write(aldeas,
         dsn="acdip_pop_project.gpkg",
         layer="aldeas",
         delete_layer=T)
```

```{r}
aldeas <- st_read(dsn="acdip_pop_project.gpkg",
                  layer="aldeas")
```


---

# Government Data

## Y2002 Census Places

```{r eval=F}
gov_cp <- st_read("data_official/lugares_poblados_gtm/lugares_poblados_gtm.shp")

# Subset Peten
gov_cp <- gov_cp %>%
  filter(departamen=="PETEN")
```

## Y2002 Census Populations

```{r}
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Ignored Population Census Data

# unclear what the shp version is, weird columns, eg DE_5_A_9
# gov_pop <- read_xls("data_official/INE_CENSO_2002_LUGAR_POBLADO/POB_TOTAL_EDADshp.xls")

# Note, POB_7_14_AÑOS_RAMA_ACTIVIDAD goes into some detail we don't need.
# gov_pop3 <- read_xls("data_official/INE_CENSO_2002_LUGAR_POBLADO/POB_7_14A_OS_RAMA_ACTIVIDAD.xls")

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Metadata: descriptor_pob_edad.dbf
# For: POB_TOTAL_EDAD

meta_pop <- read.dbf("data_official/descriptores/descriptor_pob_edad.dbf",
                     as.is=T)
# note issues with character encoding.
# meta_pop

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Data: POB_TOTAL_EDAD.xls
# Note: "data_official/POB_TOTAL_EDAD.xls" is identical
gov_pop <- read_xls("data_official/INE_CENSO_2002_LUGAR_POBLADO/POB_TOTAL_EDAD.xls")

# Subset Peten
gov_pop <- gov_pop %>%
  filter(DEPARTAMEN=="PETEN")

# # Remove GROUPS (Age Stratification)
# gov_pop <- gov_pop %>%
#   select(-contains("GRUPO"))

# fix typo in MUNICIPIO
gov_pop[gov_pop$MUNICIPIO=="LALIBERTAD", "MUNICIPIO"] <- "LA LIBERTAD"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Data: pob_edad_mun_censo02_INE.xls
# Note: I believe this is aggregated by municipio
gov_pop_aggr <- read_xls("data_official/pob_edad_mun_censo02_INE.xls")

# Subset Peten
gov_pop_aggr <- gov_pop_aggr %>%
  filter(DEPARTAMEN=="Peten")

# # Remove GROUPS (Age Stratification)
# gov_pop_aggr2 <- gov_pop_aggr %>%
#   select(-contains("GRUPO"))



gov_pop %>%
  select(-contains("GRUPO")) %>%
  group_by(MUNICIPIO) %>%
  summarize(
    summed_pop_total = sum(POBTOT),
    summed_hombres = sum(HOMBRES),
    summed_mujeres = sum(MUJERES),
    summed_urbana = sum(URBANA),
    summed_rural = sum(RURAL)
  ) %>%
  arrange(MUNICIPIO)
```

```{r}
gov_pop_aggr %>%
  arrange(MUNICIPIO)
```


### Comparing Aggregate Population Numbers in Municipios

```{r}
tmp1 <- gov_pop %>%
  select(-contains("GRUPO")) %>%
  group_by(MUNICIPIO) %>%
  summarize(summed_pop_total=sum(POBTOT),
            summed_hombres=sum(HOMBRES),
            summed_mujeres=sum(MUJERES),
            summed_urbana=sum(URBANA),
            summed_rural=sum(RURAL)
            ) %>%
  arrange(MUNICIPIO)
tmp1 <- tmp1[2:ncol(tmp1)]
```


```{r}
tmp2 <- gov_pop_aggr %>%
  select(-contains("GRUPO"), -c(AREA, OID, DEPTO, DEPARTAMEN, CODIGO)) %>%
  arrange(MUNICIPIO)
tmp2 <- tmp2[2:ncol(tmp2)]
```

```{r}
# Difference between Summed Municipios and Municipios
tmp1 - tmp2
```

```{r}
rm(tmp1, tmp2)  # gc
```


## Population Projections

```{r eval=F}
# read
gov_pp <- read_xls("data_official/proyecciones_pob_2000_2010_municipios_INE(1).xls")

# Subset Peten
gov_pp <- gov_pp %>%
  filter(DEPARTAMEN=="Peten")

gov_pp
```

The `AREA` is km^2. It is not exactly the same as the areas of derived from the IGN. This data is on the level of municipio. It is unclear what the columns N3:N13 actually represent (presumably the years). Its entirely possible that the file itself is misnamed, as the last census was in 2002, so it makes sense that this might be the projection based on 2003 onwards.

### Explore growth equation

```{r}
names(gov_pp)
```


```{r}
gov_pp2 <- gov_pp %>%
  group_by(MUNICIPIO) %>%
  select(-c(CODIGO, DEPTO, DEPARTAMEN, AREA, OID)) %>%
  gather(key="year", value="pop", ...=sapply(3:13, function(x) paste0("N", x))) %>%
  ungroup() %>%
  mutate(year=as.numeric(str_replace(year, "N", ""))) %>%
  arrange(MUNICIPIO)

gov_pp2 <- split(gov_pp2, gov_pp2$MUNICIPIO)
gov_pp2 <- lapply(gov_pp2, function(x){
  lm(pop ~ year, data=x)$coefficients
})

# populations in Y2002... is wrong!
tmp <- sapply(gov_pp2, function(x){
  unname(x[1] + (x[2] * 2))
})

tmp <- data.frame(MUNICIPIO=toupper(names(tmp)), pop_est2002=unname(tmp), stringsAsFactors = F)

gov_pop_aggr %>%
  arrange(MUNICIPIO) %>%
  mutate(MUNICIPIO = toupper(MUNICIPIO)) %>%
  select(
    -contains("GRUPO"),
    -c(
      CODIGO,
      DEPTO,
      DEPARTAMEN,
      AREA,
      HOMBRES,
      MUJERES,
      URBANA,
      RURAL,
      OID
    )
  ) %>%
  left_join(tmp, by = "MUNICIPIO")
```


## Vectors: Municipios

```{r eval=F}
gov_m <- st_read("data_official/municipios_gtm/municipios_GTM.shp")

# Subset Peten
gov_m <- gov_m %>%
  filter(Departamen=="Petén")
```

```{r eval=F}
# Sample Map
ggplot() +
  geom_sf(data=gov_m,
          mapping=aes(),
          color="#ff66ff",
          fill=NA,
          lwd=0.3)
```

