---
title: "Kriging"
author: "Kenji Tomari"
date: "3/4/2019"
output:
  html_document:
    theme: journal
    toc: yes
    toc_depth: '2'
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
libs <- c(
  "sp",
  "gstat",
  "sf",
  "tidyverse",
  "readxl",
  "rgeos"
)
suppressPackageStartupMessages({
  libs <- lapply(libs, require, character.only = TRUE)
})
rm(libs)
library(variosig)
```

```{r eval=F}
detach("package:gstat", unload=TRUE)
detach("package:geostatsp", unload=TRUE)
```


# Introduction

Here we follow this [tutorial](https://rpubs.com/nabilabd/118172).

Other Resources:

* Jef Caers @ Stanford [youtube](https://www.youtube.com/watch?v=n4oDUjjtk44&list=PLh35GyCXlQaQ1LNGWr4vCD9AGOGni8yxq&index=2)

We want to interpolate population values for unknown points

# Data

```{r}
gov_lp_inclusive <- st_read(
  dsn="acdip_pop_project.gpkg",
  layer="gov_lp_inclusive"
)

control_points <- gov_lp_inclusive %>%
  filter(POBTOT.y > 0) 

spdf <- as(control_points, "Spatial")
```

# Data Exploration

```{r}
hist(log(control_points$POBTOT.y))
```

# Find Smallest Distances to each Road Type

## Data (OSM)

```{r}
hwy_primary <- st_read(
  dsn="acdip_pop_project.gpkg",
  layer="osm_hwy_primary"
)
hwy_secondary <- st_read(
  dsn="acdip_pop_project.gpkg",
  layer="osm_hwy_secondary"
)
hwy_tertiary <- st_read(
  dsn="acdip_pop_project.gpkg",
  layer="osm_hwy_tertiary"
)

hwy_primary <- hwy_primary %>%
  st_transform(crs=st_crs(control_points)) %>%
  as("Spatial")

hwy_secondary <- hwy_secondary %>%
  st_transform(crs=st_crs(control_points)) %>%
  as("Spatial")
  
hwy_tertiary <- hwy_tertiary %>%
  st_transform(crs=st_crs(control_points)) %>%
  as("Spatial")
```

## Distances

```{r}
shortest_primary <- shortest_secondary <- shortest_tertiary <-
  numeric(nrow(spdf))

for(i in seq_len(nrow(spdf))) {
    shortest_primary[i] <- gDistance(spdf[i,], hwy_primary)
    shortest_secondary[i] <- gDistance(spdf[i,], hwy_secondary)
    shortest_tertiary[i] <- gDistance(spdf[i,], hwy_tertiary)
}

df <- data.frame(
  shortest_primary,
  shortest_secondary,
  shortest_tertiary
)

df %>%
  mutate(mean_dist = (shortest_primary +shortest_secondary+ shortest_tertiary)/3)

df$shortest_overall <- apply(df, MARGIN = 1, function(x){
  min(x["shortest_primary"], x["shortest_secondary"], x["shortest_tertiary"])
})

df$shortest_roadtype <- apply(df, MARGIN = 1, function(x){
  if(x["shortest_primary"] < x["shortest_secondary"] &
     x["shortest_primary"] < x["shortest_tertiary"]) return("primary")
  
  if(x["shortest_secondary"] < x["shortest_primary"] &
     x["shortest_secondary"] < x["shortest_tertiary"]) return("secondary")
  
  if(x["shortest_tertiary"] < x["shortest_primary"] &
     x["shortest_tertiary"] < x["shortest_secondary"]) return("tertiary")

})

spdf@data$dist_primary <- shortest_primary
spdf@data$dist_secondary <- shortest_secondary
spdf@data$dist_tertiary <- shortest_tertiary
spdf@data$shortest_overall <- df$shortest_overall
spdf@data$shortest_overall_log <- log(df$shortest_overall)
spdf@data$shortest_overall_sq <- (df$shortest_overall)^2
spdf@data$shortest_roadtype <- df$shortest_roadtype

rm(shortest_primary, shortest_secondary, shortest_tertiary)

spdf@data
```

```{r}
hist(spdf@data$shortest_overall_log)
```


# Variogram without Regressors

```{r eval=F}
lzn.vgm <- variogram(object=POBTOT.y~1, 
                     data=spdf,
                     cutoff=10000,# units are meters
                     cloud=T) 
plot(lzn.vgm)
```


# Variogram with Regressors (no log)

# Variogram using gstat

```{r}
# create gstat object (for universal kriging, so multivariate)
g <- gstat(
  id = "ln.pop",
  formula = sqrt(POBTOT.y) ~ 1,
  data = spdf,
  maxdist = 3000,
  nmax=10,
  nmin=8
)
# get variogram 
tmp <- variogram(g, cloud=T ,cutoff=3000)
plot(tmp)
```



```{r}
# create gstat object
g <- gstat(
  id = "ln.pop",
  formula = POBTOT.y ~ 1,
  data = spdf,
  maxdist = 5000,
  nmax=10,
  nmin=3
)
# get variogram 
vario <- variogram(g, cutoff=10000)
png(filename="plot_attempted_semivariogram.png",
    width=1000,
    height=700)
plot(vario, cex=4, pch=20)
```

```{r}
fit.vario <- fit.variogram(vario, model=vgm(psill=3000,
                                          model="Gau",
                                          range=3000,
                                          cutoff=3000))
plot(fit.vario, cutoff=3000)
```


```{r}
suppressMessages(gc()) 
lzn.vgm <- variogram(object=POBTOT.y ~ 1 +shortest_roadtype, 
                     data=spdf,
                     cutoff=10000,# units are meters
                     cloud=T) 
plot(lzn.vgm)
```


```{r}
suppressMessages(gc()) 
lzn.vgm <- variogram(object=POBTOT.y ~ 1 + shortest_overall_sq, 
                     locations = coordinates(spdf),
                     width=10000,
                     data=spdf,
                     cutoff=10000,  # units are meters
                     cloud=T,
                     map=T) 
plot(lzn.vgm)
```


```{r}
suppressMessages(gc()) 
lzn.vgm <- variogram(object=POBTOT.y ~ dist_primary + dist_secondary + dist_tertiary, 
                     data=spdf,
                     cutoff=10000,# units are meters
                     cloud=T) 
plot(lzn.vgm)
```

# Variogram + regressors (log)

```{r}
gc()
lzn.vgm <- variogram(object=log(POBTOT.y) ~ shortest_overall_log, 
                     data=spdf,
                     cutoff=10000,  # units are meters
                     cloud=T) 
plot(lzn.vgm)
```

```{r}
gc()
lzn.vgm <- variogram(object=log(POBTOT.y) ~ dist_primary + dist_secondary + dist_tertiary, 
                     data=spdf,
                     cutoff=10000,  # units are meters
                     cloud=T) 
plot(lzn.vgm)
```


# Variogram (alternate transform of Y)

```{r}
lzn.vgm <- variogram(object=sqrt(POBTOT.y)~1, 
                     data=spdf,
                     cutoff=5000,  # units are meters
                     cloud=T) 
plot(lzn.vgm)
```

```{r}
lzn.vgm <- variogram(object=(POBTOT.y^2)~1, 
                     data=spdf,
                     cutoff=5000,  # units are meters
                     cloud=T) 
plot(lzn.vgm)
```

```{r}
lzn.vgm <- variogram(object=log10(POBTOT.y)~1, 
                     data=spdf,
                     cutoff=5000,  # units are meters
                     cloud=T) 
plot(lzn.vgm)
```

# USING GEOSTATSP

```{r eval=F}
geo_df <- as.geodata(spdf)
spdf_variog <-  geostatsp::variog(geodata=spdf, 
                                  formula=POBTOT.y ~ 1, 
                                  option="cloud", 
                                  max.dist=3000)
spdf_variog
plot(spdf_variog)
```

# Significance Test

```{r}
# Significance
# library(variosig)
lzn.vgm <- gstat::variogram(object=log(POBTOT.y)~1, 
                     data=spdf)

# varioEnv <- envelope(lzn.vgm, 
#                      data = spdf,
#                      formula = log(POBTOT.y)~1, 
#                      nsim = 499)
# saveRDS(varioEnv, "varioEnv")
varioEnv <- readRDS("varioEnv")

png(filename="mc_semivariogram.png",
    width=1000,
    height=800)
envplot(varioEnv)
```

